---
title: "Session 6 Live Code"
author: "Carolyn Koehn"
format: html
---

This code and a few more examples can also be found on the session 6 slides.

### Access vector data:

```{r}
#| eval: false

vector.data <- sf::st_read("/opt/data/data/assignment01/cejst_nw.shp")
```
```{r}
#| include: false

vector.data <- sf::st_read("C:/Users/carolynkoehn/Documents/HES505_Fall_2024/data/2023/assignment01/cejst_nw.shp")
```

### Check the CRS:

- input
- proj4string
- wkt

```{r}
sf::st_crs(vector.data)$input
sf::st_crs(vector.data)$proj4string
sf::st_crs(vector.data)$wkt
```

### Read in raster data:

```{r}
#| eval: false

raster.data <- terra::rast("/opt/data/data/assignment01/wildfire_hazard_agg.tif")
```
```{r}
#| include: false

raster.data <- terra::rast("C:/Users/carolynkoehn/Documents/HES505_Fall_2024/data/2023/assignment01/wildfire_hazard_agg.tif")
```

### Check raster CRS:

```{r}
terra::crs(raster.data, describe=TRUE, proj=TRUE)
```

### Guess the CRS:

```{r}
library(sf)
#library(mapview)
locations <- data.frame(
  X = c(1200822.97857801, 1205015.51644983, 1202297.44383987, 1205877.68696743, 
        1194763.21511923, 1195463.42403192, 1199836.01037452, 1207081.96500368, 
        1201924.15986897),
  Y = c(1246476.31475063, 1248612.72571423, 1241479.45996392, 1243898.58428024, 
        1246033.7550009, 1241827.7730307, 1234691.50899912, 1251125.67808482, 
        1252188.4333016),
  id = 1:9
)

locations_sf <- st_as_sf(locations, coords = c("X", "Y"))
```

We stopped this example short because of confusion with the `mapview` library. Feel free to try the code from the slides on your own time!

### Plot vector data:

```{r}
plot(st_geometry(vector.data))
```

### Re-project a vector:

```{r}
vector.data.proj <- vector.data %>%
  st_transform(., crs = 3083)

st_crs(vector.data.proj)$input

plot(st_geometry(vector.data.proj))
```

### Plot a raster:

```{r}
library(terra)

plot(raster.data)
```


### Changing the CRS of a raster:

```{r}
raster.data.proj <- project(raster.data, "epsg:3083")

crs(raster.data.proj, describe=TRUE)

plot(raster.data.proj)
```

### Project based on another dataset:

```{r}
vector.data.proj.raster <- vector.data %>%
  st_transform(., crs = crs(raster.data))

st_crs(vector.data.proj.raster)$input
```

### Manually create polygon:

```{r}
outer = matrix(c(0,0,10,0,10,10,0,10,0,0),ncol=2, byrow=TRUE)
hole1 = matrix(c(1,1,1,2,2,2,2,1,1,1),ncol=2, byrow=TRUE)
hole2 = matrix(c(5,5,5,6,6,6,6,5,5,5),ncol=2, byrow=TRUE)
coords = list(outer, hole1, hole2)
pl1 = st_polygon(coords)
```

### Check polygon validity

```{r}
st_is_valid(vector.data)

table(st_is_valid(vector.data))
```

