---
title: "Session 4 Live Code"
author: "Carolyn Koehn"
format: html
---

## Reading in the data

Read spreadsheet into R:

```{r, eval=FALSE}
library(sf)
library(tidyverse)

file.to.read <- read_csv(file = "/opt/data/data/assignment01/landmarks_ID.csv", col_names = TRUE, col_type = NULL, na = c("", NA))

file.as.sf <- st_as_sf(file.to.read, coords = c("longitude", "lattitude"), crs=4326)
```
```{r, echo=FALSE}
library(sf)
library(tidyverse)

file.to.read <- read_csv(file = "C:/Users/carolynkoehn/Documents/HES505_Fall_2024/data/2023/assignment01/landmarks_ID.csv", col_names = TRUE, col_type = NULL, na = c("", NA))

file.as.sf <- st_as_sf(file.to.read, coords = c("longitude", "lattitude"), crs=4326)
```


Read in a shapefile:

```{r, eval=FALSE}
shapefile.inR <- read_sf(dsn = "/opt/data/data/assignment01/cejst_nw.shp")
```
```{r, echo=FALSE}
shapefile.inR <- read_sf(dsn = "C:/Users/carolynkoehn/Documents/HES505_Fall_2024/data/2023/assignment01/cejst_nw.shp")
```

Read in a raster:

```{r, eval=FALSE}
library(terra)
raster.inR <- rast(x = "/opt/data/data/assignment01/wildfire_hazard_agg.tif")
```
```{r, echo=FALSE}
library(terra)
raster.inR <- rast(x = "C:/Users/carolynkoehn/Documents/HES505_Fall_2024/data/2023/assignment01/wildfire_hazard_agg.tif")
```

## Introducing yourself to vector data

Structure of the data:

```{r}
str(shapefile.inR)
```

Number of rows and columns:

```{r}
nrow(shapefile.inR)
ncol(shapefile.inR)
```

Column names:

```{r}
colnames(shapefile.inR)
```

First few rows of data:

```{r}
head(file.as.sf, n = 6)
```

Simple plotting:

```{r}
plot(st_geometry(shapefile.inR))
```
```{r}
plot(shapefile.inR["AGE_10"])
```

CRS of vector:

```{r}
st_crs(shapefile.inR)
```


## Introducing yourself to raster data

Describe raster before reading it in:

```{r, eval=FALSE}
describe("/opt/data/data/assignment01/wildfire_hazard_agg.tif")
```
```{r, echo=FALSE}
describe("C:/Users/carolynkoehn/Documents/HES505_Fall_2024/data/2023/assignment01/wildfire_hazard_agg.tif")
```

Basic object info:

```{r}
raster.inR
```

Summary of values:

```{r}
# Summary of all values
summary(values(raster.inR))

# Summary of some values
summary(raster.inR)
```

Basic plot:

```{r}
plot(raster.inR)
#Change color of NA
plot(raster.inR, colNA = "black")
```

CRS of raster:

```{r}
crs(raster.inR)

# You can also use st_crs!
st_crs(raster.inR)
```

